\bf Tearing Mode Dynamics in Tokamak Plasmas

### 2021-11-24 Documentation/TearingModeTheory.tex
Latest version

Equations~(\ref{e188}) and (\ref{e189}) yield
\begin{align}\label{e198}
(1+2\,q_s^{\,2})\,\frac{d\alpha_p}{dt} + \left(\frac{1}{\tau_{\theta\,s}}+\frac{j_{1\,p}^{\,2}}{\tau_{M}}\right)\alpha_p
&= \frac{m^{2}\,[y_p(r_s/a)]^{\,2}}{\tau_A^{\,2}\,\epsilon_a^{\,2}\,[J_2(j_{1\,p})]^{\,2}}\,
{\rm Im}({\mit\Delta\hat{\Psi}_s}\,\hat{\mit\Psi}_s^{\,\ast}),\\[0.5ex]
\frac{d\beta_p}{dt} + \frac{j_{0\,p}^{\,2}}{\tau_{M}}\,\beta_p
&= \frac{n^{2}\,[z_p(r_s/a)]^{\,2}}{\tau_A^{\,2}\,[J_1(j_{0\,p})]^{\,2}}\,
{\rm Im}({\mit\Delta\hat{\Psi}_s}\,\hat{\mit\Psi}_s^{\,\ast}).\label{e199}
\end{align}
Here, 
\begin{equation}
\tau_{M} =\frac{\rho_s\,a^{\,2}}{\mu_{\perp\,s}},
\end{equation} 
is the {\em momentum confinement time}, 
\begin{equation}
\tau_A = \left(\frac{\mu_0\,\rho_s\,a^{\,2}}{B_z^{\,2}}\right)^{1/2}
\end{equation}
 is the {\em Alfv\'{e}n time}, 
 \begin{equation}
 \epsilon_a=\frac{a}{R_0}\ll 1
 \end{equation}
 is the {\em inverse aspect-ratio}\/ of the plasma, and
 \begin{align}
{\mit\Delta\hat{\Psi}}_s &=\frac{{\mit\Delta\Psi}_s}{R_0\,B_z},\\[0.5ex]
 \hat{\mit\Psi}_s &=\frac{{\mit\Psi}_s}{R_0\,B_z}.
 \end{align}
  Moreover,
use has been made of Equations~(\ref{e157}), (\ref{e158}), and (\ref{e190})--(\ref{e197}).

@@

A tearing mode is a modified shear-Alfv\'{e}n wave (Hazeltine \& Meiss 1985), and has very little connection with a compressible-Alfv\'{e}n wave. For instance, the resonance condition, ${\bf k}\cdot{\bf B}=0$, which determines the position of the resonant
surface, is identical to the resonance condition for a  shear-Alfv\'{e}n wave, but differs substantially from that of a  compressible-Alfv\'{e}n wave (Fitzpatrick 2015). Hence, one very effective way of simplifying the resonant layer equations is to remove the
physics of  compressible-Alfv\'{e}n waves from them altogether, in the process converting a 
drift-magnetohydrodynamical\footnote{The  `drift' appellation  merely indicates that we are taking diamagnetic flows into account. However, this necessarily entails treating the ion and electron flow separately. In other words, a
drift-MHD model is also a `two-fluid' model. }(MHD) response model  
into a so-called {\em reduced}\/ drift-MHD response model (Strauss 1976; Hazeltine, Kotschenreuther \& Morrison 1985). The aim of
this section is to describe the reduction process (which essentially boils down to arguing that the divergence of the
plasma flow is small), and to derive the reduced drift-MHD equations that will subsequently be used to model the resonant
response of the plasma.

\subsection{Drift-MHD Equations}
 It is helpful to define
the following parameters:
\begin{align}
n_0 &= n(r_s),\\[0.5ex]
p_0&= p(r_s),\label{e210}\\[0.5ex]
\eta_e &=\left.\frac{d\ln T_e}{d\ln n}\right|_{r=r_s},\label{e211}\\[0.5ex]
\eta_i &= \left.\frac{d\ln T_i}{d\ln n}\right|_{r=r_s},\\[0.5ex]
\tau &= \left(\frac{T_e}{T_i}\right)_{r=r_s}\left(\frac{1+\eta_e}{1+\eta_i}\right),\label{e213}
\end{align}
where $n(r)$, $p(r)$, $T_e(r)$, and $T_i(r)$ refer to density, pressure, and temperature profiles that are unperturbed by the tearing mode.

@@

\subsection{Normalization Scheme}
Let $l$ be a typical variation lengthscale in the resonant layer. The {\em Alfv\'{e}n velocity}, which is the typical phase
velocity of compressible Alfv\'{e}n waves, is defined
\begin{equation}\label{e226a}
V_A = \frac{B_z}{\sqrt{\mu_0\,n_0\,m_i}}.
\end{equation}
It is helpful to define the  {\em collisionless ion skin-depth}: 
\begin{equation}
d_i = \left(\frac{m_i}{n_0\,e^{\,2}\,\mu_0}\right)^{1/2}.
\end{equation}
Note that $d_i=c/\omega_{p\,i}$, where $c$ is the velocity of light in vacuum, and $\omega_{p\,i}$ is the ion plasma
frequency (Fitzpatrick 2015).

@@

\subsection{Reduction Process} 
All variables in the resonant layer are assumed to be functions of 
$\hat{x}=(r-r_s)/l$,
 and 
 \begin{equation}
 \zeta=m\,\theta-n\,\varphi
 \end{equation}
 only. 
Let ${\bf n} = (0,\,\epsilon/q_s,\,1)$, where $\epsilon=r/R_0$ and 
\begin{equation}
q_s=q(r_s)= \frac{m}{n}.
\end{equation}
 It follows that ${\bf n}\cdot\nabla A=0$ for any $A(\hat{x},\zeta)$. 
Let us write (Fitzpatrick \& Waelbroeck 2009)
\begin{equation}\label{e238}
\hat{\bf B} \simeq (1+ \delta b)\,{\bf n} + \hat{\nabla}\hat{\psi}\times {\bf n},
\end{equation}
which automatically satisfies Equation~(\ref{e230}). Note that $\hat{\psi}=\psi/(r_s\,B_z)$, where $\psi$ is defined in Equation~(\ref{e93}).
 Here, $\delta b(\hat{x},\zeta)$ and $\hat{\psi}(\hat{x},\zeta)$ are assumed to be small
compared to unity, which merely indicates that the perturbed magnetic field due to the tearing mode is small compared to the
equilibrium magnetic field.  We can write
\begin{equation}
\hat{\bf E} \simeq \left(\hat{E}_\parallel- \frac{\partial \hat{\psi}}{\partial \hat{t}}\right)\!{\bf n} + \hat{\nabla}\!\left(\frac{\partial\chi}{\partial \hat{t}}\right)\times {\bf n} + \hat{\nabla}\phi,
\end{equation}
which automatically satisfies Equation~(\ref{e231}) provided that
\begin{equation}
\hat{\nabla}^2\chi =\delta b.
\end{equation}
Here, $\hat{E}_\parallel = E_\parallel/(B_z\,V_A)$, where $E_\parallel$ represents the constant inductive electric field that drives the equilibrium plasma current in the vicinity of the resonant surface. Moreover, $\partial/\partial \hat{t}$ and $\phi(\hat{x},\zeta)$
are 
assumed to be first order in small quantities, whereas $\hat{E}_\parallel$ is assumed to be second order. 
It follows that
\begin{align}
\hat{\bf V}_E &\simeq \hat{\nabla}\phi\times {\bf n},\\[0.5ex]
\hat{\bf V} &\simeq \hat{V}_{\parallel\,i}\,{\bf n}+  \hat{\nabla}\phi\times {\bf n}+\hat{\nabla}{\mit\Xi},
\end{align}
where $\hat{V}_{\parallel\,i}(\hat{x},\zeta)$ is assumed to be first order in small quantities, whereas ${\mit\Xi}(\hat{x},\zeta)$ is assumed to be
second order. 
We can write
\begin{equation}
\hat{p}= \frac{\beta}{{\mit\Gamma}} + \delta p,
\end{equation}
where 
\begin{equation}\label{e244}
\beta= {\mit\Gamma}\,\hat{p}_0= \frac{{\mit\Gamma}\,p_0\,\mu_0}{B_z^{\,2}}.
\end{equation}
 Here, $\delta p(\hat{x},\zeta)$ is assumed to be first order in small
quantities. 
It follows that
\begin{align}
\hat{\bf V}_\ast &\simeq \hat{d}_i\,{\bf n}\times \hat{\nabla} \delta p,\\[0.5ex]
\hat{\bf V}_i&\simeq  \hat{V}_{\parallel\,i} \,{\bf n}+ \hat{\nabla}\!\left(\phi - \frac{\hat{d}_i\,\delta p}{1+\tau}\right)\times {\bf n}.
\end{align}

@@

\subsection{Four-Field Model}\label{fourf}
Let us define $V=\hat{d}_i\,\hat{V}_{\parallel\,i}$ and $N=-\hat{d}_i\,\delta p$. Equations~(\ref{eJ}), (\ref{e252})--(\ref{e254}), (\ref{e256}), (\ref{e258})--(\ref{e260}), (\ref{e262}), and (\ref{e263}) yield the following closed set of equations (Fitzpatrick \& Waelbroeck 2005):
\begin{align}\label{e264}
\frac{\partial\hat{\psi}}{\partial\hat{t}}&= [\phi,\hat{\psi}] -\left(\frac{\tau}{1+\tau}\right)(1+\lambda_e)\,[N,\hat{\psi}]
+\hat{\eta}_\parallel\,J + \hat{E}_\parallel,\\[0.5ex]
\frac{\partial U}{\partial \hat{t}}&= [\phi,U] - \frac{1}{2\,(1+\tau)}\left(\hat{\nabla}^2[\phi,N] + [\hat{\nabla}^2\phi,N] + [\hat{\nabla}^2 N,\phi]\right) + [J,\hat{\psi}] \nonumber\\[0.5ex]&\phantom{=}+\hat{\chi}_\varphi  \,\hat{\nabla}^4\!\left(\phi + \frac{N}{1+\tau}\right),\label{e265} \\[0.5ex]
\frac{\partial N}{\partial \hat{t}}&= [\phi,N] +c_\beta^{\,2}\,[V,\hat{\psi}] +\hat{d}_\beta^{\,2}\,[J,\hat{\psi}] + \hat{D}_\parallel\,[[N,\hat{\psi}],\hat{\psi}]
+ \hat{D}_\perp\,\hat{\nabla}_\perp^{\,2}N,\label{e266}\\[0.5ex]
\frac{\partial V}{\partial\hat{t}}&= [\phi,V] +[N,\hat{\psi}] + \hat{\chi}_\varphi\,\hat{\nabla}^2 V,\label{e267}\\[0.5ex]
J &=-\frac{2}{q_s\,\hat{R}_0}+\hat{\nabla}^2\hat{\psi},\\[0.5ex]
U &=\hat{\nabla}^2 \phi.\label{e269}
\end{align}
Here, the quantity 
\begin{equation}\label{e270}
d_\beta = c_\beta\,d_i 
\end{equation}
 is usually referred to as the {\em ion sound radius}, and $\hat{d}_\beta =d_\beta/l$. 
Moreover,
\begin{equation}
 \lambda_e = 0.71\left(\frac{\eta_e}{1+\eta_e}\right),
\end{equation}
and
\begin{align}
\hat{D}_\parallel&=\frac{D_{\parallel}}{l\,V_A}= \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\parallel\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\parallel\,i}\right),\\[0.5ex]
\hat{D}_\perp &=\frac{D_{\perp}}{l\,V_A}= c_\beta^{\,2}\,\hat{\eta}_\perp +  \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\perp\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\perp\,i}\right).
\end{align}

@@

Let us ignore the term $c_\beta^{\,2}\,d\bar{V}/dp$ in Equation~(\ref{e316}). This approximation can
be justified {\em a posteriori}. It is equivalent to the neglect of the contribution of ion parallel dynamics to the linear  plasma
response in the resonant layer, and effectively decouples Equation~(\ref{e317}) from Equations~(\ref{e314})--(\ref{e316}) (Waelbroeck, et al.\ 2012). Equations~(\ref{e314})--(\ref{e316}) reduce to (Cole \& Fitzpatrick 2006)
\begin{equation}\label{e320}
\frac{d}{dp}\!\left[G(p)\,\frac{dY}{dp}\right] - \frac{A(p)}{B(p)}\,p^{2}\,Y = 0,
\end{equation}
where
\begin{align}
G(p)&= \frac{p^2}{-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e] + p^2},\\[0.5ex]
 A(p) &= -(Q-Q_E)\,(Q-Q_E-Q_i)-{\rm i}\,(Q-Q_E-Q_i)\,(P_\varphi+P_\perp)\,p^2 + P_\varphi\,P_\perp\,p^4,\\[0.5ex]
 B(p) &= 
-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e] +\{P_\perp - {\rm i}\,(1+\lambda_e)\,(Q-Q_E-Q_i)\,D^2\}p^2\nonumber\\[0.5ex]
&\phantom{=}+[(1+\lambda_e)+1/\tau]\,P_\varphi\,D^2\,p^4,
\end{align}
and $Y=\bar{\phi} - (1+\lambda_e)\,\bar{N}$. The boundary conditions are that $Y(p)$ is bounded as $p\rightarrow\infty$, and
\begin{equation}\label{e325}
Y(p)\rightarrow Y_0\!\left[\frac{\hat{\mit\Delta}}{\pi\,p} + 1+ {\cal O}(p)\right]
\end{equation}
as $p\rightarrow 0$. In the following, we shall assume that $|Q|\sim |Q_E|\sim |Q_e|\sim |Q_i|$, $P_\varphi\sim P_\perp \sim P$, $\tau\sim {\cal O}(1)$, and $\lambda_e\sim{\cal O}(1)$,  for the
sake of simplicity.

\subsection{Constant-$\psi$ Approximation}\label{s5.6}
Let us suppose that there are two layers in $p$ space. In the small-$p$ layer, suppose that Equation~(\ref{e320}) reduces to
\begin{equation}
\frac{d}{dp}\!\left[\frac{p^2}{-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e] + p^2}\,\frac{dY}{dp}\right]\simeq 0
\end{equation}
when $p\sim Q^{1/2}$. Integrating directly, we find that
\begin{equation}
Y(p) \simeq Y_0 \left[\frac{\hat{\mit\Delta}}{\pi}\left(\frac{1}{p} + \frac{p}{{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e]}\right) + 1 + {\cal O}(p^2)\right]
\end{equation}
for $p\lesssim {\cal O}(Q^{1/2})$, where use has been made of Equation~(\ref{e325}). The two-layer  approximation is
equivalent to the well-known {\em constant-$\psi$ approximation}\/ (Furth, Killeen \& Rosenbluth 1963).

In the large-$p$ layer, for  $p\gg {\cal O}(Q^{1/2})$, we obtain 
\begin{equation}\label{e329}
\frac{d^{\,2} Y}{dp^2} - \frac{A(p)}{B(p)}\,p^2\,Y\simeq 0,
\end{equation}
with $Y$ bounded as $p\rightarrow \infty$. Asymptotic matching to the small-$p$ layer yields the boundary
condition 
\begin{equation}\label{e330}
Y(p)\simeq Y_0\left\{1+ \frac{\hat{\mit\Delta}}{\pi}\,\frac{p}{{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e]}+ {\cal O}(p^2)\right\}
\end{equation}
as $p\rightarrow 0$.

In the various constant-$\psi$ linear response regimes considered in Section~\ref{scp}, Equation~(\ref{e329}) reduces to an
equation of the form
\begin{equation}\label{e330a}
\frac{d^{\,2}Y}{dp^2}- C\,p^m\,Y\simeq 0,
\end{equation}
where $m$ is real and non-negative, and $C$ is a complex constant. Let $Y=\sqrt{p}\,Z$ and $q=\sqrt{C}\,p^n/n$, where
$n=(m+2)/2$. The previous equation transforms into a modified Bessel equation of general order, 
\begin{equation}
q^2\,\frac{d^{\,2}Z}{dq^2} + q\,\frac{dZ}{dq} - (q^2+\nu^2)\,Z =0,
\end{equation}
where $\nu=1/(m+2)$. The solution that is bounded as $q\rightarrow \infty$ has the small-$q$
expansion
\begin{equation}\label{e332}
K_\nu(z)= \frac{1}{{\mit\Gamma}(1-\nu)}\left(\frac{2}{q}\right)^\nu - \frac{1}{{\mit\Gamma}(1+\nu)}\left(\frac{q}{2}\right)^\nu + {\cal O}(q^{2-\nu}),
\end{equation}
where ${\mit\Gamma}(z)$ is a gamma function. 
A comparison of this expression with Equation~(\ref{e330}) reveals that
\begin{equation}
\hat{\mit\Delta} = \pi\,(-{\rm i}\,\left[Q-Q_E-(1+\lambda_e)\,Q_e\right])\, \frac{{\mit\Gamma}(1-\nu)}{{\mit\Gamma}(\nu)}\,\nu^{2\nu-1}\,C^{\,\nu}.
\end{equation}
Note, finally, that $p_\ast\sim |C|^{-\nu}$, where $p_\ast$ denotes the width of the large-$p$ layer in $p$ space. This width must be larger than $Q^{1/2}$ (i.e., the width of the small-$p$ layer) in order for the constant-$\psi$ approximation to hold.

>>>>

Equations~(\ref{e188}) and (\ref{e189}) yield
\begin{align}\label{e198}
(1+2\,q_s^{\,2})\,\frac{d\alpha_p}{dt} + \left(\frac{1}{\tau_{\theta\,s}}+\frac{j_{1\,p}^{\,2}}{\tau_{M}}\right)\alpha_p
&= \frac{m^{2}\,[y_p(r_s/a)]^{\,2}}{\tau_A^{\,2}\,\epsilon_a^{\,2}\,[J_2(j_{1\,p})]^{\,2}}\,
{\rm Im}({\mit\Delta\hat{\Psi}_s}\,\hat{\mit\Psi}_s^{\,\ast}),\\[0.5ex]
\frac{d\beta_p}{dt} + \frac{j_{0\,p}^{\,2}}{\tau_{M}}\,\beta_p
&= \frac{n^{2}\,[z_p(r_s/a)]^{\,2}}{\tau_A^{\,2}\,[J_1(j_{0\,p})]^{\,2}}\,
{\rm Im}({\mit\Delta\hat{\Psi}_s}\,\hat{\mit\Psi}_s^{\,\ast}).\label{e199}
\end{align}
Here, 
\begin{equation}
\tau_{M} =\frac{\rho_s\,a^{\,2}}{\mu_{\perp\,s}},
\end{equation} 
is the {\em momentum confinement time}, 
\begin{equation}
\tau_A = \left(\frac{\mu_0\,\rho_s\,a^{\,2}}{B_z^{\,2}}\right)^{1/2}
\end{equation}
 is the {\em Alfv\'{e}n time}, 
 \begin{equation}
 \epsilon_a=\frac{a}{R_0}\ll 1
 \end{equation}
 is the {\em inverse aspect-ratio}\/ of the plasma, and
 \begin{align}
{\mit\Delta\hat{\Psi}}_s &=\frac{{\mit\Delta\Psi}_s}{R_0\,B_z},\label{e204}\\[0.5ex]
 \hat{\mit\Psi}_s &=\frac{{\mit\Psi}_s}{R_0\,B_z}.
 \end{align}
  Moreover,
use has been made of Equations~(\ref{e157}), (\ref{e158}), and (\ref{e190})--(\ref{e197}).

@@

A tearing mode is a modified shear-Alfv\'{e}n wave (Hazeltine \& Meiss 1985), and has very little connection with a compressible-Alfv\'{e}n wave. For instance, the resonance condition, ${\bf k}\cdot{\bf B}=0$, which determines the position of the resonant
surface, is identical to the resonance condition for a  shear-Alfv\'{e}n wave, but differs substantially from that of a  compressible-Alfv\'{e}n wave (Fitzpatrick 2015). Hence, one very effective way of simplifying the resonant layer equations is to remove the
physics of  compressible-Alfv\'{e}n waves from them altogether, in the process converting a 
drift-magnetohydrodynamical\footnote{The  `drift' appellation  merely indicates that we are taking diamagnetic flows into account. However, this necessarily entails treating the ion and electron flows separately. In other words, a
drift-MHD model is,  by definition, also a `two-fluid' model. }(MHD) response model  
into a so-called {\em reduced}\/ drift-MHD response model (Strauss 1976; Hazeltine, Kotschenreuther \& Morrison 1985). The aim of
this section is to describe the reduction process (which essentially boils down to arguing that the divergence of the
plasma flow is small), and to derive the reduced drift-MHD equations that will subsequently be used to model the resonant
response of the plasma.

\subsection{Drift-MHD Equations}\label{s4.2}
 It is helpful to define
the following parameters:
\begin{align}
n_0 &= n(r_s),\\[0.5ex]
p_0&= p(r_s),\label{e210}\\[0.5ex]
\eta_e &=\left.\frac{d\ln T_e}{d\ln n}\right|_{r=r_s},\label{e211}\\[0.5ex]
\eta_i &= \left.\frac{d\ln T_i}{d\ln n}\right|_{r=r_s},\\[0.5ex]
\tau &= \left(\frac{T_e}{T_i}\right)_{r=r_s}\left(\frac{1+\eta_e}{1+\eta_i}\right),\label{e213}
\end{align}
where $n(r)$, $p(r)$, $T_e(r)$, and $T_i(r)$ refer to density, pressure, and temperature profiles that are unperturbed by the tearing mode.

@@

\subsection{Normalization Scheme}\label{s4.3}
Let $l$ be a typical variation lengthscale in the resonant layer. The {\em Alfv\'{e}n velocity}, which is the typical phase
velocity of compressible Alfv\'{e}n waves, is defined
\begin{equation}\label{e226a}
V_A = \frac{B_z}{\sqrt{\mu_0\,n_0\,m_i}}.
\end{equation}
It is helpful to define the  {\em collisionless ion skin-depth}: 
\begin{equation}
d_i = \left(\frac{m_i}{n_0\,e^{\,2}\,\mu_0}\right)^{1/2}.
\end{equation}
Note that $d_i=c/\omega_{p\,i}$, where $c$ is the velocity of light in vacuum, and $\omega_{p\,i}$ is the ion plasma
frequency (Fitzpatrick 2015).

@@

\subsection{Reduction Process} 
All variables in the resonant layer are assumed to be functions of 
$\hat{x}=(r-r_s)/l$,
 \begin{equation}
 \zeta=m\,\theta-n\,\varphi,
 \end{equation}
 and $\hat{t}$, 
 only. 
Let ${\bf n} = (0,\,\epsilon/q_s,\,1)$, where $\epsilon=r/R_0$ and 
\begin{equation}
q_s=q(r_s)= \frac{m}{n}.
\end{equation}
 It follows that ${\bf n}\cdot\nabla A=0$ for any $A(\hat{x},\zeta,\hat{t})$. 
Let us write (Fitzpatrick \& Waelbroeck 2009)
\begin{equation}\label{e238}
\hat{\bf B} \simeq (1+ \delta b)\,{\bf n} + \hat{\nabla}\hat{\psi}\times {\bf n},
\end{equation}
which automatically satisfies Equation~(\ref{e230}). Note that 
\begin{equation}\label{e247a}
\hat{\psi}=\frac{\psi}{r_s\,B_z},
\end{equation}
 where $\psi$ is defined in Equation~(\ref{e93}).
 Here, $\delta b(\hat{x},\zeta,
 \hat{t})$ and $\hat{\psi}(\hat{x},\zeta,\hat{t})$ are assumed to be small
compared to unity, which merely indicates that the perturbed magnetic field due to the tearing mode is small compared to the
equilibrium magnetic field.  We can write
\begin{equation}
\hat{\bf E} \simeq \left(\hat{E}_\parallel- \frac{\partial \hat{\psi}}{\partial \hat{t}}\right)\!{\bf n} + \hat{\nabla}\!\left(\frac{\partial\chi}{\partial \hat{t}}\right)\times {\bf n} + \hat{\nabla}\phi,
\end{equation}
which automatically satisfies Equation~(\ref{e231}) provided that
\begin{equation}
\hat{\nabla}^2\chi =\delta b.
\end{equation}
Here, $\hat{E}_\parallel = E_\parallel/(B_z\,V_A)$, where $E_\parallel$ represents the constant inductive parallel electric field that drives the equilibrium plasma current in the vicinity of the resonant surface. Moreover, $\partial/\partial \hat{t}$ and $\phi(\hat{x},\zeta,\hat{t})$
are 
assumed to be first order in small quantities, whereas $\hat{E}_\parallel$ is assumed to be second order. 
It follows that
\begin{align}
\hat{\bf V}_E &\simeq \hat{\nabla}\phi\times {\bf n},\\[0.5ex]
\hat{\bf V} &\simeq \hat{V}_{\parallel\,i}\,{\bf n}+  \hat{\nabla}\phi\times {\bf n}+\hat{\nabla}{\mit\Xi},
\end{align}
where $\hat{V}_{\parallel\,i}(\hat{x},\zeta,\hat{t})$ is assumed to be first order in small quantities, whereas ${\mit\Xi}(\hat{x},\zeta,\hat{t})$ is assumed to be
second order. 
We can write
\begin{equation}
\hat{p}= \frac{\beta}{{\mit\Gamma}} + \delta p,
\end{equation}
where 
\begin{equation}\label{e244}
\beta= {\mit\Gamma}\,\hat{p}_0= \frac{{\mit\Gamma}\,p_0\,\mu_0}{B_z^{\,2}}.
\end{equation}
 Here, $\delta p(\hat{x},\zeta,\hat{t})$ is assumed to be first order in small
quantities. 
It follows that
\begin{align}
\hat{\bf V}_\ast &\simeq \hat{d}_i\,{\bf n}\times \hat{\nabla} \delta p,\\[0.5ex]
\hat{\bf V}_i&\simeq  \hat{V}_{\parallel\,i} \,{\bf n}+ \hat{\nabla}\!\left(\phi - \frac{\hat{d}_i\,\delta p}{1+\tau}\right)\times {\bf n}.
\end{align}

@@

\subsection{Four-Field Model}\label{fourf}
Let us define $V=\hat{d}_i\,\hat{V}_{\parallel\,i}$ and $N=-\hat{d}_i\,\delta p$. Equations~(\ref{eJ}), (\ref{e252})--(\ref{e254}), (\ref{e256}), (\ref{e258})--(\ref{e260}), (\ref{e262}), and (\ref{e263}) yield the following closed set of equations (Fitzpatrick \& Waelbroeck 2005):
\begin{align}\label{e264}
\frac{\partial\hat{\psi}}{\partial\hat{t}}&= [\phi,\hat{\psi}] -\left(\frac{\tau}{1+\tau}\right)(1+\lambda_e)\,[N,\hat{\psi}]
+\hat{\eta}_\parallel\,J + \hat{E}_\parallel,\\[0.5ex]
\frac{\partial U}{\partial \hat{t}}&= [\phi,U] - \frac{1}{2\,(1+\tau)}\left(\hat{\nabla}^2[\phi,N] + [\hat{\nabla}^2\phi,N] + [\hat{\nabla}^2 N,\phi]\right) + [J,\hat{\psi}] \nonumber\\[0.5ex]&\phantom{=}+\hat{\chi}_\varphi  \,\hat{\nabla}^4\!\left(\phi + \frac{N}{1+\tau}\right),\label{e265} \\[0.5ex]
\frac{\partial N}{\partial \hat{t}}&= [\phi,N] +c_\beta^{\,2}\,[V,\hat{\psi}] +\hat{d}_\beta^{\,2}\,[J,\hat{\psi}] + \hat{D}_\parallel\,[[N,\hat{\psi}],\hat{\psi}]
+ \hat{D}_\perp\,\hat{\nabla}_\perp^{\,2}N,\label{e266}\\[0.5ex]
\frac{\partial V}{\partial\hat{t}}&= [\phi,V] +[N,\hat{\psi}] + \hat{\chi}_\varphi\,\hat{\nabla}^2 V,\label{e267}\\[0.5ex]
J &=-\frac{2}{q_s\,\hat{R}_0}+\hat{\nabla}^2\hat{\psi},\\[0.5ex]
U &=\hat{\nabla}^2 \phi.\label{e269}
\end{align}
Here, the quantity 
\begin{equation}\label{e270}
d_\beta = c_\beta\,d_i 
\end{equation}
 is usually referred to as the {\em ion sound radius}, and $\hat{d}_\beta =d_\beta/l$. 
Moreover,
\begin{equation}
 \lambda_e = 0.71\left(\frac{\eta_e}{1+\eta_e}\right),
\end{equation}
and
\begin{align}
\hat{D}_\parallel&\equiv\frac{D_{\parallel}}{l\,V_A}= \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\parallel\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\parallel\,i}\right),\\[0.5ex]
\hat{D}_\perp &\equiv \frac{D_{\perp}}{l\,V_A}= c_\beta^{\,2}\,\hat{\eta}_\perp +  \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\perp\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\perp\,i}\right).
\end{align}

@@

Let us ignore the term $c_\beta^{\,2}\,d\bar{V}/dp$ in Equation~(\ref{e316}). This approximation can
be justified {\em a posteriori}. It is equivalent to the neglect of the contribution of ion parallel dynamics to the linear  plasma
response in the resonant layer, and effectively decouples Equation~(\ref{e317}) from Equations~(\ref{e314})--(\ref{e316}) (Waelbroeck, et al.\ 2012). Equations~(\ref{e314})--(\ref{e316}) reduce to (Cole \& Fitzpatrick 2006)
\begin{equation}\label{e320}
\frac{d}{dp}\!\left[G(p)\,\frac{dY_e}{dp}\right] - \frac{A(p)}{B(p)}\,p^{2}\,Y_e = 0,
\end{equation}
where
\begin{align}
G(p)&= \frac{p^2}{-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e] + p^2},\\[0.5ex]
 A(p) &= -(Q-Q_E)\,(Q-Q_E-Q_i)-{\rm i}\,(Q-Q_E-Q_i)\,(P_\varphi+P_\perp)\,p^2 + P_\varphi\,P_\perp\,p^4,\\[0.5ex]
 B(p) &= 
-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e] +\{P_\perp - {\rm i}\,(1+\lambda_e)\,(Q-Q_E-Q_i)\,D^2\}p^2\nonumber\\[0.5ex]
&\phantom{=}+[(1+\lambda_e)+1/\tau]\,P_\varphi\,D^2\,p^4,
\end{align}
and $Y_e=\bar{\phi} - (1+\lambda_e)\,\bar{N}$. The boundary conditions are that $Y_e(p)$ is bounded as $p\rightarrow\infty$, and
\begin{equation}\label{e325}
Y_e(p)\rightarrow Y_0\!\left[\frac{\hat{\mit\Delta}}{\pi\,p} + 1+ {\cal O}(p)\right]
\end{equation}
as $p\rightarrow 0$. In the following, we shall assume that $|Q|\sim |Q_E|\sim |Q_e|\sim |Q_i|$, $P_\varphi\sim P_\perp \sim P$, $\tau\sim {\cal O}(1)$, and $\lambda_e\sim{\cal O}(1)$,  for the
sake of simplicity.

\subsection{Constant-$\psi$ Approximation}\label{s5.6}
Let us suppose that there are two layers in $p$ space. In the small-$p$ layer, suppose that Equation~(\ref{e320}) reduces to
\begin{equation}
\frac{d}{dp}\!\left[\frac{p^2}{-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e] + p^2}\,\frac{dY_e}{dp}\right]\simeq 0
\end{equation}
when $p\sim Q^{1/2}$. Integrating directly, we find that
\begin{equation}
Y_e(p) \simeq Y_0 \left[\frac{\hat{\mit\Delta}}{\pi}\left(\frac{1}{p} + \frac{p}{{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e]}\right) + 1 + {\cal O}(p^2)\right]
\end{equation}
for $p\lesssim {\cal O}(Q^{1/2})$, where use has been made of Equation~(\ref{e325}). The two-layer  approximation is
equivalent to the well-known {\em constant-$\psi$ approximation}\/ (Furth, Killeen \& Rosenbluth 1963).

In the large-$p$ layer, for  $p\gg {\cal O}(Q^{1/2})$, we obtain 
\begin{equation}\label{e329}
\frac{d^{\,2} Y_e}{dp^2} - \frac{A(p)}{B(p)}\,p^2\,Y_e\simeq 0,
\end{equation}
with $Y_e$ bounded as $p\rightarrow \infty$. Asymptotic matching to the small-$p$ layer yields the boundary
condition 
\begin{equation}\label{e330}
Y_e(p)\simeq Y_0\left\{1+ \frac{\hat{\mit\Delta}}{\pi}\,\frac{p}{{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e]}+ {\cal O}(p^2)\right\}
\end{equation}
as $p\rightarrow 0$.
